\bf \texttt{forestecology} package for modeling interspecies competition between trees

### 2021-02-12 etc/paper/forestecology.tex

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(forestecology)}

\CommentTok{# Load tidyverse packages:}
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(lubridate)}

\CommentTok{# Load spatial packages:}
\KeywordTok{library}\NormalTok{(blockCV)}
\KeywordTok{library}\NormalTok{(sf)}
\KeywordTok{library}\NormalTok{(sfheaders)}

\CommentTok{# Load other packages:}
\KeywordTok{library}\NormalTok{(snakecase)}
\KeywordTok{library}\NormalTok{(yardstick)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{(bw_census_}\DecValTok{2008}\NormalTok{, bw_census_}\DecValTok{2014}\NormalTok{, bw_species)}

\CommentTok{# Append additional species data}
\NormalTok{bw_census_}\DecValTok{2008}\NormalTok{ <-}\StringTok{ }\NormalTok{bw_census_}\DecValTok{2008} \OperatorTok{%>%}
\StringTok{  }\KeywordTok{left_join}\NormalTok{(bw_species, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(genus, species, latin))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_}\DecValTok{2013}\NormalTok{ <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =}\NormalTok{ ExactDate,}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh),}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh }\OperatorTok{/}\StringTok{ }\DecValTok{10}\NormalTok{)}

\NormalTok{scbi_}\DecValTok{2018}\NormalTok{ <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =}\NormalTok{ ExactDate,}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh),}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh }\OperatorTok{/}\StringTok{ }\DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bw_census_}\DecValTok{2014}\NormalTok{ <-}\StringTok{ }\NormalTok{bw_census_}\DecValTok{2014} \OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{))}

\NormalTok{bw_growth_df <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(bw_census_}\DecValTok{2008}\NormalTok{, bw_census_}\DecValTok{2014}\NormalTok{, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# Convert all variables denoting species to factors}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{sp =}\NormalTok{ sp }\OperatorTok{%>%}\StringTok{ }\KeywordTok{to_any_case}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{as.factor}\NormalTok{(),}
    \DataTypeTok{species =}\NormalTok{ sp,}
    \DataTypeTok{family =} \KeywordTok{as.factor}\NormalTok{(family),}
    \DataTypeTok{trait_group =} \KeywordTok{as.factor}\NormalTok{(trait_group)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# Drop unnecessary variables}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\NormalTok{stemID)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_growth_df <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(scbi_}\DecValTok{2013}\NormalTok{, scbi_}\DecValTok{2018}\NormalTok{, }\StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# Convert all variables denoting species to factors}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{as.factor}\NormalTok{(sp))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{max_dist <-}\StringTok{ }\FloatTok{7.5}
\NormalTok{cv_fold_size <-}\StringTok{ }\DecValTok{100}
\end{Highlighting}
\end{Shaded}

The first constant is \texttt{max\_dist} which defines the maximum
distance for a tree's competitive neighborhood. Trees within this
distance of each other are assumed to compete while those farther than
this distance apart do not. Put differently, all trees within
\texttt{max\_dist} of a focal tree will be considered its competitors
(see below). Other studies have estimated the value of
\texttt{max\_dist}; we use an average of estimated values
\citet{canham_neighborhood_2004}, \citet{uriarte_spatially_2004},
\citet{tatsumi2013}, \citet{canham_neighborhood_2006}.

Furthermore, \texttt{max\_dist} will define the size of all buffers
considered, which will be encoded as a binary variable \texttt{buffer}
as computed by the \texttt{add\_buffer\_variable()} function. This
function takes as input the main \texttt{growth\_df} data frame, the
\texttt{size} of the buffer which we set as \texttt{max\_dist}, and the
boundary of the study region encoded as a simple features polygon
\citet{pebesma_simple_2018}. DESCRIBE SF PACKAGE. In the Big Woods
example below we will use a pre-loaded simple features polygon while for
the SCBI example we present example code on how to manually construct
one.

@@

First, we indicate which trees are part of the buffer. This necessitates
information about the study region boundary. In this case, we use a
\texttt{sf\_polygon} object \texttt{bw\_study\_region} which comes
pre-loaded in the \texttt{forestecology} packages. After loading
\texttt{bw\_study\_region}, we illustrate the results of the
\texttt{add\_buffer\_variable()} function in Figure
\ref{fig:bw-define-buffer}. Trees on the periphery denote with lighter
colors are part of the buffer and will not be considered as ``focal''
trees of interest going forward; they will only be considered as
competitor trees.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{(bw_study_region)}

\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }
                      \DataTypeTok{region =}\NormalTok{ bw_study_region)}

\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }
          \KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{76}\NormalTok{)}
\NormalTok{bw_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
  \DataTypeTok{speciesData =}\NormalTok{ bw_growth_df, }\DataTypeTok{theRange =}\NormalTok{ cv_fold_size, }\DataTypeTok{k =} \DecValTok{28}\NormalTok{, }\DataTypeTok{xOffset =} \FloatTok{0.5}\NormalTok{, }
  \DataTypeTok{yOffset =} \DecValTok{0}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{foldID)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Visualize grid. Why does fold 19 repeat?}
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_as_sf}\NormalTok{()) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }
          \KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{factor}\NormalTok{(foldID)), }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_as_sf}\NormalTok{(), }
               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ folds))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\NormalTok{foldID }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{19}\NormalTok{, }\DecValTok{23}\NormalTok{, }\DecValTok{21}\NormalTok{, }\DecValTok{17}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{19}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =} \KeywordTok{as.character}\NormalTok{(foldID))}
\end{Highlighting}
\end{Shaded}

Separately, we save the spatial cross-validation grid as an
\texttt{sf\_polygon} object \texttt{bw\_cv\_grid}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bw_cv_grid <-}\StringTok{ }\NormalTok{bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{st_as_sf}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_study_region <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}

\NormalTok{scbi_growth_df <-}\StringTok{ }\NormalTok{scbi_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }
                      \DataTypeTok{region =}\NormalTok{ scbi_study_region)}

\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_growth_df, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

Next we define \texttt{focal\_vs\_comp} data frames which connects each
focal tree in the \texttt{growth\_df} data frames to the trees in its
competitive neighborhood range as defined by the \texttt{max\_dist}
constant. So for example, if \texttt{growth\_df} consisted of two focal
trees with two and three neighbors with \texttt{max\_dist} respectively,
\texttt{focal\_vs\_comp} would be a data frame of 5 rows connecting each
focal tree to it's competitors. The \texttt{create\_focal\_vs\_comp()}
function makes this connection taking as inputs the \texttt{growth\_df}
data frame; the \texttt{max\_dist} constant defining competitive range;
\texttt{cv\_grid\_sf}, giving the cross validation grid; and the
\texttt{id} variable.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ bw_cv_grid, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{scbi_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ scbi_cv_grid, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots_bw[[}\StringTok{"beta_0"}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

>>>>

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tidyverse)}
\KeywordTok{library}\NormalTok{(lubridate)}
\KeywordTok{library}\NormalTok{(sf)}
\KeywordTok{library}\NormalTok{(forestecology)}
\KeywordTok{library}\NormalTok{(blockCV)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{census_}\DecValTok{2008}\NormalTok{_bw <-}\StringTok{ }\NormalTok{census_}\DecValTok{2008}\NormalTok{_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{left_join}\NormalTok{(species_bw, }\DataTypeTok{by =} \StringTok{"sp"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(}\OperatorTok{-}\KeywordTok{c}\NormalTok{(genus, species, latin))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =}\NormalTok{ ExactDate }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mdy}\NormalTok{(),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh }\OperatorTok{/}\StringTok{ }\DecValTok{10}\NormalTok{)}

\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(treeID, stemID, sp, ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =}\NormalTok{ ExactDate }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mdy}\NormalTok{(),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{300}\NormalTok{, }\KeywordTok{between}\NormalTok{(gy, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{dbh =}\NormalTok{ dbh }\OperatorTok{/}\StringTok{ }\DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_bw <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2008}\NormalTok{_bw }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2014}\NormalTok{_bw }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
    \DataTypeTok{id =} \StringTok{"treeID"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{      }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{to_any_case}\NormalTok{(sp) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{()),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{comp_dist <-}\StringTok{ }\FloatTok{7.5}
\NormalTok{cv_fold_size <-}\StringTok{ }\DecValTok{100}
\end{Highlighting}
\end{Shaded}

The first constant is \texttt{comp\_dist} which defines the maximum
distance for a tree's competitive neighborhood. Trees within this
distance of each other are assumed to compete while those farther than
this distance apart do not. Put differently, all trees within
\texttt{comp\_dist} of a focal tree will be considered its competitors
(see below). Other studies have estimated the value of
\texttt{comp\_dist}; we use an average of estimated values
\citet{canham_neighborhood_2004}, \citet{uriarte_spatially_2004},
\citet{tatsumi2013}, \citet{canham_neighborhood_2006}.

Furthermore, \texttt{comp\_dist} will define the size of all buffers
considered, which will be encoded as a binary variable \texttt{buffer}
as computed by the \texttt{add\_buffer\_variable()} function. This
function takes as input the main \texttt{growth\_df} data frame, the
\texttt{size} of the buffer which we set as \texttt{comp\_dist}, and the
boundary of the study region encoded as a simple features polygon
\citet{pebesma_simple_2018}. DESCRIBE SF PACKAGE. In the Big Woods
example below we will use a pre-loaded simple features polygon while for
the SCBI example we present example code on how to manually construct
one.

@@

First, we indicate which trees are part of the buffer. This necessitates
information about the study region boundary. In this case, we use a
\texttt{sf\_polygon} object \texttt{study\_region\_bw} which comes
pre-loaded in the \texttt{forestecology} packages. After loading
\texttt{study\_region\_bw}, we illustrate the results of the
\texttt{add\_buffer\_variable()} function in Figure
\ref{fig:bw-define-buffer}. Trees on the periphery denote with lighter
colors are part of the buffer and will not be considered as ``focal''
trees of interest going forward; they will only be considered as
competitor trees.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{(study_region_bw)}

\NormalTok{growth_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_bw)}

\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_bw }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{76}\NormalTok{)}
\NormalTok{bw_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
  \DataTypeTok{speciesData =}\NormalTok{ growth_bw, }\DataTypeTok{theRange =}\NormalTok{ cv_fold_size, }\DataTypeTok{k =} \DecValTok{28}\NormalTok{, }\DataTypeTok{xOffset =} \FloatTok{0.5}\NormalTok{, }
  \DataTypeTok{yOffset =} \DecValTok{0}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{foldID)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Visualize grid. Why does fold 19 repeat?}
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_as_sf}\NormalTok{()) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_bw }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }
          \KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{factor}\NormalTok{(foldID)), }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_as_sf}\NormalTok{(), }
               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ folds))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\NormalTok{foldID }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{19}\NormalTok{, }\DecValTok{23}\NormalTok{, }\DecValTok{21}\NormalTok{, }\DecValTok{17}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{19}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =} \KeywordTok{factor}\NormalTok{(foldID))}
\end{Highlighting}
\end{Shaded}

Separately, we save the spatial cross-validation grid as an
\texttt{sf\_polygon} object \texttt{blocks\_bw}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{blocks_bw <-}\StringTok{ }\NormalTok{bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{st_as_sf}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{600}\NormalTok{, }\DecValTok{300}\NormalTok{)}
\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}

\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}

\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ growth_scbi, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

Next we define \texttt{focal\_vs\_comp} data frames which connects each
focal tree in the \texttt{growth\_df} data frames to the trees in its
competitive neighborhood range as defined by the \texttt{comp\_dist}
constant. So for example, if \texttt{growth\_df} consisted of two focal
trees with two and three neighbors with \texttt{comp\_dist}
respectively, \texttt{focal\_vs\_comp} would be a data frame of 5 rows
connecting each focal tree to it's competitors. The
\texttt{create\_focal\_vs\_comp()} function makes this connection taking
as inputs the \texttt{growth\_df} data frame; the \texttt{comp\_dist}
constant defining competitive range; \texttt{cv\_grid\_sf}, giving the
cross validation grid; and the \texttt{id} variable.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{growth_bw }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_bw, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(comp_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{plot1 <-}\StringTok{ }\KeywordTok{autoplot}\NormalTok{(comp_bayes_lm_bw, }\DataTypeTok{type =} \StringTok{"intercepts"}\NormalTok{)}
\NormalTok{plot1}
\end{Highlighting}
\end{Shaded}
